\LARGE \bf The Battle for Filter Supremacy: A Retrospective: The Movie: The Game: The Paper

### 2014-12-29 Report/project_report.tex
Added MSCKF and IMU data for plotting
Final commit for the night. 2 new sets of plots added to the pdf
Fixed plot axes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Extended Kalman Filter (EKF)} \label{subsec:ekf}
In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of features by recursively updating the state as new measurements become available.
%Features are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
%In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sliding Window Filter (SWF)} \label{subsec:slidingwindow}
In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
\begin{equation}
(\Vec H^T \Vec T ^{-1} \Vec H ) \delta \Vec{x}^*  = - \Vec{H}^T\Vec{T}^{-1}\Vec{e}(\bar{\Vec{x}})
\end{equation}
\begin{equation}
\delta\Vec{x} := \bbm \delta \Vec x_0 & \hdots & \delta \Vec x_K & \delta \Vec f_1 & \hdots & \delta \Vec f_N\ebm ^T,
\end{equation}
where $N$ is the number of unique features observed in the window.
The matrix $\Vec{H}$ is given by (showing non-zero blocks only)
\begin{equation}
\mathbf{H} = \left[\begin{array}{cccccc}
 -\Vec H_{\Vec x, 1} & \Vec 1 & &  & &   \\ 
 & -\Vec G_{\Vec x, 1} &  &  &  &   -\Vec G_{\Vec f, 1} \\ \hline
 & -\Vec H_{\Vec x, 2} & \Vec 1 & & &    \\ 
 & & -\Vec G_{\Vec x, 2} &  & &  -\Vec G_{\Vec f, 2}   \\ \hline
 &  & \ddots & \ddots & & \vdots  \\  \hline
  & & & -\Vec H_{\Vec x, K} & \Vec 1 & \\ 
 & & & & -\Vec G_{\Vec x, K} &  -\Vec G_{\Vec f, K}   \\  
\end{array}\right], 
\end{equation}
where
\begin{align}
\Vec G_{\Vec{x}, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}}
    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\boldsymbol{\rho}_i^{f_j G} - \Vec r_G^{IG,k} ) )^\times  \ebm ^T
\\[0.5em]
\Vec G_{\Vec x, k} &= \bbm \Vec G_{\Vec x, k}^1 & \hdots & \Vec G_{\Vec x, k}^M \ebm ^T
\\[0.5em]
\Vec G_{\Vec f, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} \Vec C_{CI} \Vec C_{IG,k}
\\[0.5em]
\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j$}}
\end{align}
with
\begin{equation}
\bar{\Vec p}_{C_k}^{f_j C_k} := \bbm x \\ y \\ z \ebm = \Vec C_{CI} \left( \Vec C_{IG,k} (\boldsymbol{\rho}_G^{f_j G} - \Vec r_G^{IG,k} )   - \boldsymbol \rho_I^{CI} \right),
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MULTI-STATE CONSTRAINT KALMAN FILTER} \label{sec:msckf}
The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the number of vehicle states in the window exceeds some preset threshold.

@@

Next, we define a vector of three parameters:
\begin{align}
\hat{\Vec{y}} &= \bbm \alpha & \beta & \gamma \ebm^T = \frac{1}{\hat{Z}^{(j)}_{C_1}} \bbm \hat{X}^{(j)}_{C_1} & \hat{Y}^{(j)}_{C_1} & 1 \ebm ^T
\end{align}
The feature position in camera frame $C_i$ can be expressed as
\begin{align}
\hat{\Vec{p}}_{C_i}^{f_j C_i} = \hat{\Vec{C}}_{i1} \hat{\Vec{p}}_{C_1}^{f_j C_1}  + \hat{\Vec{p}}_{C_i}^{C_1 C_i}
\end{align}
With our state parameters, we can use this equation to define three functions:
\begin{align}
    \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \\ h_3 (\hat{\Vec{y}}) \ebm 
    & = \hat{\Vec{C}}_{i1} \bbm \alpha \\ \beta \\ 1 \ebm + \gamma  \hat{\Vec{p}}_{C_i}^{C_1 C_i} \notag \\
    & = \bbm C_{11} \alpha + C_{12} \beta + C_{13} + \gamma \hat{x}_{C_i}^{C_1 C_i} \\[0.5em]
            C_{21} \alpha + C_{22} \beta + C_{23} + \gamma \hat{y}_{C_i}^{C_1 C_i} \\[0.5em]
            C_{31} \alpha + C_{32} \beta + C_{33} + \gamma \hat{z}_{C_i}^{C_1 C_i} 
        \ebm 
\end{align}
The camera measurement error can then be written as
\begin{align}
\Vec e(\hat{\Vec{y}}) = \hat{\Vec{z}}_i^{(j)} - \frac{1}{h_3 (\hat{\Vec{y}}) } \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \ebm
\end{align}
The least squares linear system then becomes
\begin{align}
(\Vec  E^T \Vec W ^{-1} \Vec E ) \delta \Vec{y}^*  = - \Vec{E}^T\Vec{W}^{-1}\Vec{e}(\hat{\Vec{y}})
\end{align}
where
\begin{align}
\Vec E &= \frac{\partial \Vec e}{\partial \hat{\Vec{y}}}
& &
\text{and}
&
\Vec W &= \bbm \Vec{R}_1'^{(j)} & & \\ & \ddots & \\ & & \Vec{R}_N'^{(j)} \ebm 
\end{align}
with $\Vec{R}_i'^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.

@@

By stacking the errors $\Vec{r}_i^{(j)}$, we arrive at an expression for the $2M_j \times 1$ measurement error vector of feature $f_j$ over the entire window of camera poses, where $M_j$ is the number of camera poses from which feature $f_j$ was observed:
\begin{align} \label{eq:residual_stacked}
    \Vec{r}^{(j)} &\simeq \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}^{(j)}
\end{align}
The noise vector $\Vec{n}^{(j)}$ has covariance matrix $\Vec{R}^{(j)} = \text{diag}\left\{\Vec{R}_1^{(j)}, \hdots, \Vec{R}_{M_j}^{(j)}\right\}$.

However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the error in the state.
Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly. [TALK HERE ABOUT WHAT HAPPENS IF YOU LEAVE IT IN]

In order to transform $\Vec{r}^{(j)}$ into a usable form for the EKF, we can define a semi-unitary matrix $\Vec{A}$ whose columns form the basis of the left nullspace of $\Vec{H}_{f}^{(j)}$, and project $\Vec{r}^{(j)}$ into this nullspace to obtain an error equation of the correct form:
\begin{align}
    \Vec{r}_o^{(j)} &:= \Vec{A}^T \left(\Vec{z}_i^{(j)} - \hat{\Vec{z}}_i^{(j)}\right) \\
                    &\simeq \Vec{A}^T \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{0} + \Vec{A}^T \Vec{n}^{(j)} \\
                    &=: \Vec{H}_{o}^{(j)} \widetilde{\Vec{x}} + \Vec{n}_o^{(j)}
\end{align}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Extended Kalman Filter (EKF)} \label{sec:EKF}
In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of features by recursively updating the state as new measurements become available.
%Features are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
%In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sliding Window Filter (SWF)} \label{sec:SWF}
In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
\begin{equation}
(\Vec H^T \Vec T ^{-1} \Vec H ) \delta \Vec{x}^*  = - \Vec{H}^T\Vec{T}^{-1}\Vec{e}(\bar{\Vec{x}})
\end{equation}
\begin{equation}
\delta\Vec{x} := \bbm \delta \Vec x_0 & \hdots & \delta \Vec x_K & \delta \Vec f_1 & \hdots & \delta \Vec f_N\ebm ^T,
\end{equation}
where $N$ is the number of unique features observed in the window.
The matrix $\Vec{H}$ is given by (showing non-zero blocks only)
\begin{equation}
\mathbf{H} = \left[\begin{array}{cccccc}
 -\Vec H_{\Vec x, 1} & \Vec 1 & &  & &   \\ 
 & -\Vec G_{\Vec x, 1} &  &  &  &   -\Vec G_{\Vec f, 1} \\ \hline
 & -\Vec H_{\Vec x, 2} & \Vec 1 & & &    \\ 
 & & -\Vec G_{\Vec x, 2} &  & &  -\Vec G_{\Vec f, 2}   \\ \hline
 &  & \ddots & \ddots & & \vdots  \\  \hline
  & & & -\Vec H_{\Vec x, K} & \Vec 1 & \\ 
 & & & & -\Vec G_{\Vec x, K} &  -\Vec G_{\Vec f, K}   \\  
\end{array}\right], 
\end{equation}
where
\begin{align}
\Vec G_{\Vec{x}, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}}
    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\boldsymbol{\rho}_i^{f_j G} - \Vec r_G^{IG,k} ) )^\times  \ebm ^T
\\[0.5em]
\Vec G_{\Vec x, k} &= \bbm \Vec G_{\Vec x, k}^1 & \hdots & \Vec G_{\Vec x, k}^M \ebm ^T
\\[0.5em]
\Vec G_{\Vec f, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} \Vec C_{CI} \Vec C_{IG,k}
\\[0.5em]
\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j$}}
\end{align}
with
\begin{equation}
\bar{\Vec p}_{C_k}^{f_j C_k} := \bbm x \\ y \\ z \ebm = \Vec C_{CI} \left( \Vec C_{IG,k} (\boldsymbol{\rho}_G^{f_j G} - \Vec r_G^{IG,k} )   - \boldsymbol \rho_I^{CI} \right),
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MULTI-STATE CONSTRAINT KALMAN FILTER} \label{sec:MSCKF}
The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the number of vehicle states in the window exceeds some preset threshold.

@@

Next, we define a vector of three parameters:
\begin{align}
\hat{\Vec{y}} &= \bbm \alpha & \beta & \gamma \ebm^T = \frac{1}{\hat{Z}^{(j)}_{C_1}} \bbm \hat{X}^{(j)}_{C_1} & \hat{Y}^{(j)}_{C_1} & 1 \ebm ^T
\end{align}
The feature position in camera frame $C_i$ can be expressed as
\begin{align}
\hat{\Vec{p}}_{C_i}^{f_j C_i} = \hat{\Vec{C}}_{i1} \hat{\Vec{p}}_{C_1}^{f_j C_1}  + \hat{\Vec{p}}_{C_i}^{C_1 C_i}
\end{align}
With our state parameters, we can use this equation to define three functions:
\begin{align}
    \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \\ h_3 (\hat{\Vec{y}}) \ebm 
    & = \hat{\Vec{C}}_{i1} \bbm \alpha \\ \beta \\ 1 \ebm + \gamma  \hat{\Vec{p}}_{C_i}^{C_1 C_i} \notag \\
    & = \bbm C_{11} \alpha + C_{12} \beta + C_{13} + \gamma \hat{x}_{C_i}^{C_1 C_i} \\[0.5em]
            C_{21} \alpha + C_{22} \beta + C_{23} + \gamma \hat{y}_{C_i}^{C_1 C_i} \\[0.5em]
            C_{31} \alpha + C_{32} \beta + C_{33} + \gamma \hat{z}_{C_i}^{C_1 C_i} 
        \ebm 
\end{align}
The camera measurement error can then be written as
\begin{align}
\Vec e(\hat{\Vec{y}}) = \hat{\Vec{z}}_i^{(j)} - \frac{1}{h_3 (\hat{\Vec{y}}) } \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \ebm
\end{align}
The least squares linear system then becomes
\begin{align}
(\Vec  E^T \Vec W ^{-1} \Vec E ) \delta \Vec{y}^*  = - \Vec{E}^T\Vec{W}^{-1}\Vec{e}(\hat{\Vec{y}})
\end{align}
where
\begin{align}
\Vec E &= \frac{\partial \Vec e}{\partial \hat{\Vec{y}}}
& &
\text{and}
&
\Vec W &= \text{diag}\left\{ \Vec{R}_1'^{(j)}, \hdots, \Vec{R}_N'^{(j)} \right\}
\end{align}
with $\Vec{R}_i'^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.

@@

By stacking the errors $\Vec{r}_i^{(j)}$, we arrive at an expression for the $2M_j \times 1$ measurement error vector of feature $f_j$ over the entire window of camera poses, where $M_j$ is the number of camera poses from which feature $f_j$ was observed:
\begin{align} \label{eq:residual_stacked}
    \Vec{r}^{(j)} &= \Vec{z}^{(j)} - \hat{\Vec{z}}^{(j)} \simeq \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}^{(j)}
\end{align}
The noise vector $\Vec{n}^{(j)}$ has covariance matrix $\Vec{R}^{(j)} = \text{diag}\left\{\Vec{R}_1^{(j)}, \hdots, \Vec{R}_{M_j}^{(j)}\right\}$.

However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the error in the state.
Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly.
Figure \ref{fig:nullspace} shows the effect of attempting to use this error directly in the EKF.
The correlation between the measurement error and the state causes the filter to behave very poorly after a small number of time steps.

\begin{figure}
    \centering
    \subfigure {
        \includegraphics[width=0.48\textwidth]{figs/NST_trans}
        \label{fig:nullspace:trans_err}
    }

    \subfigure {
        \includegraphics[width=0.48\textwidth]{figs/NST_rot}
        \label{fig:nullspace:rot_err}
    }

    \caption{Effect of nullspace projection on MSCKF performance}
    \label{fig:nullspace}
\end{figure}

In order to transform $\Vec{r}^{(j)}$ into a usable form for the EKF, we can define a semi-unitary matrix $\Vec{A}$ whose columns form the basis of the left nullspace of $\Vec{H}_{f}^{(j)}$, and project $\Vec{r}^{(j)}$ into this nullspace to obtain an error equation of the correct form (to first order):
\begin{align}
    \Vec{r}_o^{(j)} := \Vec{A}^T \Vec{r}^{(j)} 
                    &\simeq \Vec{A}^T \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{0} + \Vec{A}^T \Vec{n}^{(j)} \notag \\
                    &=: \Vec{H}_{o}^{(j)} \widetilde{\Vec{x}} + \Vec{n}_o^{(j)}
\end{align}

@@

\begin{figure}
    \centering
    \subfigure {
        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-500-1000-Trans}
        \label{fig:comp_1:trans_err}
    }

    \subfigure {
        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-500-1000-Rot}
        \label{fig:comp_1:rot_err}
    }

    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on the dataset interval $k\in[500,1000]$. The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature tracks lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
    \label{fig:comp_1}
\end{figure}

\begin{figure}
    \centering
    \subfigure {
        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-1215-1715-Trans}
        \label{fig:comp_2:trans_err}
    }

    \subfigure {
        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-1215-1715-Rot}
        \label{fig:comp_2:rot_err}
    }

    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on the dataset interval $k\in[1215,1715]$. The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature tracks lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
    \label{fig:comp_2}
\end{figure}
